### 2015-09-21 iclr-paper.tex
more changes. added figures for mnist

\documentclass{article} % For LaTeX2e
\usepackage{iclr-style/iclr2016_conference,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{caption}

\newcommand{\sigmoid}{\boldsymbol{\sigma}}
\newcommand{\hlang}{h^{lang}}
\newcommand{\hlangall}{\boldsymbol{h^{lang}}}
\newcommand{\hdec}{h^{dec}}
\newcommand{\henc}{h^{enc}}
\newcommand{\readop}{\mathit{read}}
\newcommand{\writeop}{\mathit{write}}
\newcommand{\encoder}{\mathit{LSTM}^{enc}}
\newcommand{\decoder}{\mathit{LSTM}^{dec}}
\newcommand{\canv}{c}
\newcommand{\lat}{z}
\newcommand{\Lat}{Z}
\newcommand{\post}{Q}
\newcommand{\prior}{P}
\newcommand{\loss}{\mathcal{L}}
\newcommand{\lloss}{\mathcal{L}^{z}}
\newcommand{\rloss}{\mathcal{L}^{x}}

\newenvironment{myfont}{\fontfamily{[pcr]}\selectfont}{\par}
\DeclareTextFontCommand{\textmyfont}{\myfont}

\setlength{\abovecaptionskip}{5pt} % Chosen fairly arbitrarily
\setlength{\belowcaptionskip}{5pt} % Chosen fairly arbitrarily

\title{Generating Images From Captions\\ With Attention}

\author{
Elman Mansimov\textsuperscript{*}, Emilio Parisotto\thanks{Equal Contribution} , Jimmy Lei Ba \& Ruslan Salakhutdinov\\
Department of Computer Science\\
University of Toronto\\
Toronto, Ontario, Canada \\
\texttt{\{emansim,eparisotto,rsalakhu\}@cs.toronto.edu}, \texttt{jimmy@psi.utoronto.ca}
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\begin{document}

\maketitle

\begin{abstract}
At the end.
\end{abstract}

\section{Introduction}

Generating realistic images from their descriptions is a very hard task that combines together two challenging problems of language modelling and image generation. Effectively, the model has to capture the semantic meaning expressed in the description and then use that knowledge to generate the pixel intensities of the image. Generating images from captions is a significantly harder problem than a reverse problem of generating captions from images which has recenly attracted a lot of attention from machine learning research community \citep{karpathy_captions}, \citep{xu_captions}, \citep{kiros_captions} and etc.

By using a sequence to sequence framework to approach the problem of image generation from captions, our model iteratively draws the patches on canvas, while attending to the relevant words in the description.

\section{Related Work}

Deep Neural Networks have achieved a remarkable performance in various tasks such as image recognition \citep{krizhevsky_imagenet}, speech transcription \citep{graves_speech} and etc. While most of the recent success has been achieved by discriminative models, the generative models have not yet enjoyed the same level of success. Most of the previous work in generative models has been focused on variants of Boltzmann Machines \citep{smolensky_rbm}, \citep{russ_dbm} and Deep Belief Networks \citep{hinton_dbn}. While these models are very powerful, each iteration of training requires a computationally costly step of MCMC to approximate an intractable normalization constant that makes it hard to scale them to large datasets.

\cite{kingma_vae} have introduced the Variational Auto-Encoder (VAE) which can be seen as a neural network with continous latent variables. The encoder is used to approximate posterior distribution and the decoder is used to stochastically reconstruct the data from latent variables. The model performs an efficient inference and learning that allows it to scale to large datasets. \cite{gregor_draw} have introduced Deep Recurrent Attention Writer (DRAW), where they have incorporated a novel differentiable attention mechanism into the VAE which significantly improved its performance as well as quality of generated samples. While most of the samples of VAE and DRAW resemble a clear structure of objects, the generated images are blurry most of the time.

Generative Adversarial Networks (GANs) \citep{goodfellow_gan} are another type of generative models that use noise-contrastive estimation \citep{gutmann_nce} to avoid calculating intractable normalization constant. The model consists of generator that generates samples from uniform distribution and discriminator that discriminates between real and generated images. Both networks are playing the game, where generator tries to produce samples that look real and discriminator tries not to be fooled by generator. Recently, \cite{denton_lapgan} have scaled those models, by training GANs at each level of Laplacian pyramid of images. While their model has generated sharp looking samples, the generated images have lacked a clear structure. Compared to other mentioned generative models, GANs are more unstable and are harder to train.

While all of the previous work has been focused on unconditional models or models conditioned on labels, to the best of our knowledge this paper is the first to introduce the generative model of images conditioned on captions.

\section{Model}

Our proposed model can be seen as a part of sequence-to-sequence framework \citep{ilya_mt}, \citep{cho_mt}, \citep{nitish_video} where captions are represented as a sequence of consecutive words and images are represented as a sequence of patches drawn on canvas over time $t=1,...,T$. Let $y$ be the input caption, consisting of $N$ words $y_{1}, y_{2}, ..., y_{n}$ and $x$ be the image corresponding to that caption.

\subsection{Encoder}

The encoder is a deterministic Bidirectional LSTM that encodes the variable size sentences into the vector representation $s$. Bidirectional LSTM consists of one Forward LSTM and Backward LSTM which combine information from past and future respectively. The Forward LSTM computes the sequence of forward hidden states $[\overrightarrow{h}^{lang}_{1}, \overrightarrow{h}^{lang}_{2}, ..., \overrightarrow{h}^{lang}_{N}]$ , whereas the Backward LSTM computes the sequence of backward hidden states $[\overleftarrow{h}^{lang}_{1}, \overleftarrow{h}^{lang}_{2}, ..., \overleftarrow{h}^{lang}_{N}]$. Then these hidden states are concatenated together into the sequence $[\hlang_{1}, \hlang_{2}, ..., \hlang_{N}]$, where $\hlang_{n} = [\overrightarrow{h}^{lang}_{n}, \overleftarrow{h}^{lang}_{n}], 1\leq n\leq N$.

The final representation of the sentence is calculated as follows:
\begin{align}
s_{t} = \alpha_{1}\hlang_{1} + \alpha_{2}\hlang_{2} + ... + \alpha_{N}\hlang_{N}
\end{align}
where $\hlang_{0}$ is initialized to the learned bias.
Setting $\alpha_{1...N}$ to $\frac{1}{N}$ turns the encoder into the vanilla model introduced in \citep{cho_mt} without the attention. We will describe how $\alpha_{1...N}$ are calculated in the next section.

\subsection{Decoder}

The decoder is a DRAW model, which is a stochastic recurrent neural network that consists of Inference LSTM that infers the distribution of latent variables of image $x$ given $y$ and then the Generator LSTM that uses the inferred latent variables in order to reconstruct the image $x$ given $y$. Formally, the decoder iteratively computes the following equations for $t=1,...,T$
\begin{align}
\label{eq:x_hat}
\hat{x}_t &= x-\sigmoid(\canv_{t-1})\\
\label{eq:read}
r_t &= \readop(x_t, \hat{x}_t, \hdec_{t-1})\\
\henc_t &= \encoder(\henc_{t-1}, [r_t, \hdec_{t-1}])\\
\lat_t &\sim \post(\Lat_t|\henc_t)\\
\hdec_t &= \decoder(\hdec_{t-1}, z_t, s_{t-1})\\
s_{t} &= align(\hdec_{t-1}, \hlangall)\\
\label{eq:write}
\canv_t &= \canv_{t-1} + \writeop(\hdec_t)
\end{align}

where $\readop$ and $\writeop$ are the same attention operators as in \citep{gregor_draw}, $\hlangall = [\hlang_{1}, \hlang_{2}, ..., \hlang_{n}]$ and $\canv_{0}, \hdec_{0}, \henc_{0}$ are initialized to the learned biases.

The $align$ function introduced by \cite{bahdanau_mt} is used to compute probabilities $\alpha_{1...n}$ by first computing the energies $e_{t,0}, e_{t,1}, ..., e_{t,n}$ and then rescaling them to the probability distribution $\alpha_{1}, \alpha_{2}, ..., \alpha_{n}$
\begin{align}
e_{tj} &= v^{T}tanh(U\hlang_{j} + W\hdec_{t} + b)\\
\alpha_{j} &= softmax(e_{tj})
\end{align}

where $softmax(e_{tj}) = \frac{exp(e_{tj})}{\sum_{j=1}^{N}exp(e_{tj})}$

\subsection{Learning}

The model is learned by the modified version of Stochastic Gradient Variation Bayes (SGVB) algorithm introduced by \cite{kingma_vae}. The model is trained to maximize the lower bound of marginal likelihood $\loss$ of the correct image $x$ given the input caption $y$. The $\loss$ is decomposed into the latent loss $\lloss$ and the reconstruction loss $\rloss$.

The reconstruction loss $\rloss$ equals to $\frac{1}{L}\sum_{l=1}^{L}(log\,p(x_{t}|y,z)$ where $L$ is the number of samples used during training, which was set to $1$ in our experiments.

The latent loss is a negative sum of Kullback--Leibler divergence terms between distribution $\post(\Lat_t|\henc_t)$ and some prior distribution ${\prior(\Lat_t)}$ over time $t=1,...,T$, which can be seen as a regularization term. Since the patches drawn on canvas over time are not independent of each other, naturally the sufficient statistics of the prior distribution at time $t$ should be dependent on the sufficient statistics of the prior distribution at time $t-1$. Therefore, instead of setting $\prior(\Lat_1), ..., \prior(\Lat_T)$ to be independent unit gaussian distributions, the mean and variance of $\prior(\Lat_t)$ depends on the $\hdec_{t-1}$, as in \citep{bachman_sdm}, where 
\begin{align}
\mu_{t}^{prior} &= tanh(W_{mu}\hdec_{t-1})\\
\sigma_{t}^{prior} &= exp(tanh(W_{\sigma}\hdec_{t-1})) 
\end{align}
Overall, the likelihood $\loss$ is calculated as follows:
\begin{align}
\loss &= -\sum_{t=1}^{T}D_{KL}(\post(\Lat_t|\henc_t,s_{t-1})\,||\,\prior(\Lat_t)) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)\\
&=
\frac{1}{2}\sum_{t=1}^{T}(1 - 2\,log\,\sigma_{t}^{prior} + 2\,log\,\sigma_{t} - \frac{exp(2\,log\,\sigma_{t}) + (\mu_{t} - \mu_{t}^{prior})^{2}}{exp(2\,log\,\sigma_{t}^{prior})}) + \frac{1}{L}\sum_{l=1}^{L}log\,p(x_{t}|y,z)
\end{align}

\subsection{Image Generation}

During the image generation step, we throw away the inference network from the decoder and instead sample from the prior distribution. Due to the blurriness of samples generated by DRAW model, we do an additional post processing step, where we use the generator of the adversarial network trained on residuals of laplacian pyramid to sharpen the generated images, similar to \citep{denton_lapgan}. By fixing the prior of adversarial generator to the mean of uniform distribution, it gets treated as a deterministic neural network which allows us to calculate the lower bound of likelihood. The reconstuction loss becomes the loss between sharpened image and correct image, whereas the latent loss stays the same. We also noticed that sampling from the mean of uniform distribution allowed us to generate much less noisy samples than by sampling from the uniform distribution itself.

\section{Experiments}
\subsection{MNIST With Captions}
As a toy experiment, we have trained our model on the two digit MNIST dataset with captions constructed on the fly. Two random digits were either placed horizontally or vertically, so that they don't overlap, on the black $60 \times 60$ background. The caption indicated the way digits were placed in the image and had the following template: \textit{the digit (1) is (2) the (3) of the digit (4) \textless eos\textgreater}. (1) and (4) were the digit numbers, (2) was either on or at, (3) was one of the adjectives top, bottom, left or right. For example, the generated images corresponding to the caption \textit{the digit seven is at the bottom of the digit five \textless eos\textgreater} as well as other captions are shown on~\ref{fig:fig1}. While most of the generative models were trained on the version of binarized MNIST dataset \citep{russ_binarizedmnist}, our model was trained directly on pixel intensities with binary cross entropy cost function.

\subsection{COCO}
TODO:
1) show table of SSI and PR curve
2) show samples unfolded over time as well as the words the model is attending to (both success and failure cases)
3) show generated samples with colour flipped, background changed etc.
4) compare samples with LAPGAN, VAE etc.
5) show tsne plot of types of images and well as topic of samples where the model works well where works badly

\subsection{CIFAR}

\begin{figure}[t]
\label{fig1}
\captionsetup{labelformat=empty}
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-4.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-6.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-9.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-13.png}
  \caption{\textit{is digit 5}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-17.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-22.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-26.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/5-7-27.png}
  \caption{\mbox{\textit{eos 7 digit}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-4.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-6.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-9.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-13.png}
  \caption{\mbox{\textit{eos digit 2}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-17.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-22.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-26.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/9-2-27.png}
  \caption{\mbox{\textit{of top digit}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-4.png}
  \caption{\mbox{\textit{eos 0 of}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-6.png}
  \caption{\mbox{\textit{eos digit 0}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-9.png}
  \caption{\mbox{\textit{eos digit 0}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-12.png}
  \caption{\mbox{\textit{eos 0 of}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-17.png}
  \caption{\mbox{\textit{1 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-22.png}
  \caption{\mbox{\textit{1 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-25.png}
  \caption{\mbox{\textit{1 digit on}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/1-0-28.png}
  \caption{\mbox{\textit{digit the eos}}}
\endminipage\hfill

\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-4.png}
  \caption{\textit{left 8 4}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-6.png}
  \caption{\textit{left 8 4}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-9.png}
  \caption{\mbox{\textit{left digit 4}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-14.png}
  \caption{\mbox{\textit{eos 8 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-18.png}
  \caption{\mbox{\textit{eos 8 digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-22.png}
  \caption{\mbox{\textit{8 digit the}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-25.png}
  \caption{\mbox{\textit{8 digit digit}}}
\endminipage\hfill
\minipage{0.1\textwidth}
  \includegraphics[width=\linewidth]{figures/4-8-28.png}
  \caption{\mbox{\textit{digit digit the}}}
\endminipage\hfill

\captionsetup{labelformat=empty}
\caption{Figure 1: Four examples show the generated images unfolded over several timesteps as well as the top three words the model attends to while generating images.}
\end{figure}

\bibliography{iclr-paper}
\bibliographystyle{iclr-style/iclr2016_conference}

\end{document}